Template notebook for validating cell type assignments for an individual library in SCPCP000015

### 2025-01-28 analyses/cell-type-ewings/template_notebooks/celltype-exploration.Rmd
add some documentation for new functions

# source in validation functions calculate_mean_markers()
validation_functions <- file.path(module_base, "scripts", "utils", "tumor-validation-helpers.R")
source(validation_functions)
```

@@

```{r, fig.height = 5, fig.width=5}
plot_faceted_umap(all_info_df, singler_lumped, legend_title = "SingleR cell types") +
  theme(strip.text = element_text(size = 8))
```

@@

```{r, fig.height = 5, fig.width=5}
plot_faceted_umap(all_info_df, cluster, legend_title = "Cluster")
```

@@

```{r, fig.height=10, fig.width=10}
expression_umap(auc_df, auc_value)
```

@@

```{r, fig.height=10, fig.width=5}
auc_columns <- colnames(all_info_df)[which(startsWith(colnames(all_info_df), "auc_"))]
cluster_density_plot(all_info_df, auc_columns, "AUC")
```

@@

```{r}
annotated_exp_heatmap(
  all_info_df, 
  exp_columns = auc_columns, 
  annotation_column = "singler_lumped", 
  cluster_column = "cluster", 
  legend_title = "AUC"
)
```

@@

expression_umap(exp_df, mean)
```

@@

```{r}
annotated_exp_heatmap(
  all_info_df, 
  exp_columns = mean_exp_columns, 
  annotation_column = "singler_lumped", 
  cluster_column = "cluster", 
  legend_title = "Mean\ngene expression"
)
```

>>>>

# source in validation functions 
# calculate_mean_markers(), plot_faceted_umap()
validation_functions <- file.path(module_base, "scripts", "utils", "tumor-validation-helpers.R")
source(validation_functions)

# source in plotting functions 
# expression_umap(), cluster_density_plot(), and annotated_exp_heatmap()
plotting_functions <- file.path(module_base, "template_notebooks", "utils", "plotting-functions.R")
source(plotting_functions)
```

@@

```{r, fig.height = 5}
plot_faceted_umap(all_info_df, singler_lumped, legend_title = "SingleR cell types") +
  theme(strip.text = element_text(size = 8))
```

@@

```{r, fig.height = 5}
plot_faceted_umap(all_info_df, cluster, legend_title = "Cluster")
```

@@

```{r, fig.height=10, fig.width=10}
expression_umap(auc_df, auc_value, geneset)
```

@@

```{r, fig.height=10}
auc_columns <- colnames(all_info_df)[which(startsWith(colnames(all_info_df), "auc_"))]
cluster_density_plot(all_info_df, auc_columns, "AUC")
```

@@

```{r}
annotated_exp_heatmap(
  all_info_df, 
  exp_columns = auc_columns, 
  cell_type_column = "singler_lumped", 
  cluster_column = "cluster", 
  legend_title = "AUC"
)
```

@@

expression_umap(exp_df, mean, geneset)
```

@@

```{r}
annotated_exp_heatmap(
  all_info_df, 
  exp_columns = mean_exp_columns, 
  cell_type_column = "singler_lumped", 
  cluster_column = "cluster", 
  legend_title = "Mean\ngene expression"
)
```
